---
title: "Course project 1"
author: "Mohanad"
date: "7/17/2020"
output: html_document
---
Here we are gonna load the data and delete the NA Values

```{r}
row_data=read.csv("./activity.csv")
data=na.omit(row_data)
data$date=as.Date(data$date)
```
let us make another data frame called "stepsbyday" and aggregate the steps by date
```{r}
stepsbyday <- aggregate(steps~date,data=data,sum)
```
let us plot the histogram 
```{r}
hist(stepsbyday$steps)
```
we will compute the mean and median of stepsbyday
```{r}
mean(stepsbyday$steps)
median(stepsbyday$steps)
``` 
we will now calculate the average by day and plot it
```{r}
averagebyday <- aggregate(steps~interval,data=data,mean)
plot(averagebyday$interval,averagebyday$steps,type='l',main="Average by day",xlab="Interval",ylab="Average of steps")
```

```{r} 
maximum=max(averagebyday$steps)
averagebyday[grepl(maximum,averagebyday$steps),]$interval
```
let us replace the number of rows with NA with the mean for 5-minute interval
```{r}
data_imputed <- row_data
for (i in 1:nrow(data_imputed)) {
  if (is.na(data_imputed$steps[i])) {
     markup<-data_imputed$interval[i]
     data_imputed$steps[i]<-averagebyday[averagebyday$interval%in%markup,]$steps
     
  }
}
head(data_imputed)
df_imputed_steps_by_day <- aggregate(steps ~ date, data_imputed, sum)
hist(df_imputed_steps_by_day$steps, main="Histogram of total number of steps per day (imputed)", 
     xlab="Total number of steps in a day")
mean(df_imputed_steps_by_day$steps)
median(df_imputed_steps_by_day$steps)
```
*Are there differences in activity patterns between weekdays and weekends?*
```{r}
data_imputed['type_of_day'] <- weekdays(as.Date(data_imputed$date))
data_imputed$type_of_day[data_imputed$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
data_imputed$type_of_day[data_imputed$type_of_day != "weekend"] <- "weekday"
data_imputed$type_of_day <- as.factor(data_imputed$type_of_day)

# calculate average steps by interval across all days
df_imputed_steps_by_interval <- aggregate(steps ~ interval + type_of_day, data_imputed, mean)
library(ggplot2)
# creat a plot
qplot(interval, 
      steps, 
      data = df_imputed_steps_by_interval, 
      type = 'l', 
      geom=c("line"),
      xlab = "Interval", 
      ylab = "Number of steps", 
      main = "") +
  facet_wrap(~ type_of_day, ncol = 1)
```

